Real Estate

### 2022-11-29 src/real_estate.Rmd
inverted income to home price column

```{r}
# add MSA_median rent data
compliedMSA_realEstate<- select_MSAs %>% 
  left_join(MSA_realEstate_data) %>% 
    # calculate rent change
  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
  # calculate income to homeprice
  mutate(house_price_2021 = as.numeric(house_price_2021)) %>% 
  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
  # calculate rent to income ratio
  # multiply monthly rent x 12 for annual rent!
  mutate(avg_rent_to_income = (avg_rent_2021*12)/median_income) %>% 
  mutate(med_rent_to_income = (median_rent2021*12)/median_income)

@@

```{r}
# normalize columns into new data table
n_MSA_realEstate <- compliedMSA_realEstate %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
  # normalize employment growth
  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
  # normalize rent to income
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income, na.rm = TRUE)) / (max(avg_rent_to_income, na.rm = TRUE) - min(avg_rent_to_income, na.rm = TRUE))) %>% 
  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
  # normalize rent change
  mutate(n_avg_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))

@@

```{r}
# invert normalized rent to income ratio so lower values (more favorable) have higher scores
n_MSA_realEstate <- n_MSA_realEstate %>% 
  mutate(n_med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
  mutate(n_avg_rent_to_income_inverted = 1 - n_avg_rent_to_income)

@@

#make real_estate_score column
n_MSA_realEstate <- n_MSA_realEstate %>% 
  select(cbsa, MSA, starts_with("n_"), ) %>% 
  mutate(real_estate_score = MSA) %>% #this is just a place holder
  mutate(RE_score_avg_rent = MSA)
```

@@

for (i in 1:size[1]) {
  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
                  n_MSA_realEstate$n_employ_change[i],
                  n_MSA_realEstate$n_avg_occupancy[i],
                  n_MSA_realEstate$n_med_rent_change[i],
                  n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
                  n_MSA_realEstate$n_income_to_homeprice[i]
                  )
  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
}

@@

```{r}

# score using average rent
for (i in 1:size[1]) {
  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
                  n_MSA_realEstate$n_employ_change[i],
                  n_MSA_realEstate$n_avg_occupancy[i],
                  n_MSA_realEstate$n_avg_rent_change[i],
                  n_MSA_realEstate$n_avg_rent_to_income_inverted[i], # compare avg rent and median rent
                  n_MSA_realEstate$n_income_to_homeprice[i]
                  )
  n_MSA_realEstate$RE_score_avg_rent[i] = weighted.mean(MSA_x_data, real_estate_weights)
}
```

>>>>

```{r}
# add MSA_median rent data
compliedMSA_realEstate<- select_MSAs %>% 
  left_join(MSA_realEstate_data) %>% 
    # calculate rent change
  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
  # calculate income to homeprice
  mutate(house_price_2021 = as.numeric(house_price_2021)) %>% 
  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
  # calculate rent to income ratio
  # multiply monthly rent x 12 for annual rent!
  mutate(med_rent_to_income = (median_rent2021*12)/median_income)

@@

```{r}
# normalize columns into new data table
n_MSA_realEstate <- compliedMSA_realEstate %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
  # normalize employment growth
  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
  # normalize rent to income
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
  # normalize rent change
  mutate(n_avg_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))

@@

```{r}
# invert normalized rent to income ratio so lower values (more favorable) have higher scores
n_MSA_realEstate <- n_MSA_realEstate %>% 
  mutate(n_med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
  mutate(n_income_to_homeprice_inverted = 1 - n_income_to_homeprice)

@@

#make real_estate_score column
n_MSA_realEstate <- n_MSA_realEstate %>% 
  select(cbsa, MSA, starts_with("n_"), ) %>% 
  mutate(real_estate_score = MSA) 
```

@@

for (i in 1:size[1]) {
  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
                  n_MSA_realEstate$n_employ_change[i],
                  n_MSA_realEstate$n_avg_occupancy[i],
                  n_MSA_realEstate$n_med_rent_change[i],
                  n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
                  n_MSA_realEstate$n_income_to_homeprice_inverted[i]
                  )
  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
}

@@
